A Classical Density-Functional Theory for Describing Water Interfaces

### 2011-06-27 paper/poster.tex
add nice little cartoon for the poster.

\definecolor{cyan}{cmyk}{0.4,0,0.3,0}
\definecolor{greenblue}{rgb}{0.7,1,.7}
\definecolor{mediumblue}{rgb}{0.7,0.7,1}
\begin{poster}{
  % Show grid to help with alignment
  grid=no,
  % Column spacing
  colspacing=1em,
  % Color style
  bgColorOne=lightyellow,
  bgColorTwo=lightestyellow,
  borderColor=mediumblue,
  headerColorOne=lighterblue,
  headerColorTwo=mediumblue,
  headerFontColor=black,
  boxColorOne=lighterblue,
  boxColorTwo=lightblue,
  % Format of textbox
  textborder=roundedleft,
  % Format of text header
  eyecatcher=yes,
  headerborder=open,
  headerheight=0.08\textheight,
  headershape=roundedright,
  headershade=shade-tb,
  headerfont=\Large\textsf, %Sans Serif
  boxshade=plain,
%  background=shade-tb,
  background=shade-tb,
  linewidth=2pt
  }
  % Eye Catcher
  { % The title is centered between eye-catcher and logo.
    \includegraphics[height=5.5em]{figs/cavitysize-060}
  }
  % Title
  {\sf %Sans Serif
  %\bf% Serif
  A Classical Density Functional for Water}
  % Authors
  {\sf %Sans Serif
  % Serif
  David Roundy\hspace{3em}
  Jessica Hughes\hspace{3em}
  Oregon State University, USA
  }
  % University logo
  {
    \includegraphics[height=5.5em]{figs/osu-logo}
  }

@@

  \headerbox{Classical DFT}{name=cdft,column=0,row=0}{ Classical
    Density-Functional Theory is a continuum empirical approach to
    modelling fluid behavior when the density is inhomogeneous---as is
    the case at interfaces.  We have developed a classical density
    functional for water, which provides a description of hydrophobic
    interactions.  }

  \headerbox{Statistical Associating Fluid
    Teory}{name=saft,column=0,below=cdft}{ The statistical associating
    fluid theory (SAFT) is a well-established approach for the
    description of the thermodynamics and phase equilibria of fluid
    systems, with particular applicability to hydrogen-bonding fluids.
    Our functional reduces in the homogeneous limit to the SAFT model
    for water developed in reference~\cite{clark2006developing}.
    \begin{center}
      \includegraphics[width=\columnwidth]{figs/pressure-with-isotherms}
    \end{center}
    This functional accurately describes the liquid-vapor phase
    diagram away from the critical point, but gives a poor description
    of the compressibility.
  }

  \headerbox{The
    Functional}{name=fitting,column=0,below=saft,above=bottom}{ We
    introduce a new classical density functional for water.  The
    Helmholtz free energy functional is based on SAFT, and is composed
    of the usual terms:
    \begin{equation}
      F[n] = F_\textit{id}[n] + F_\textit{hs}[n] + F_\textit{assoc}[n] + F_\textit{disp},
    \end{equation}
    We model the repulsive interactions using the White Bear version
    of the Fundamental-Measure Theory~(FMT)
    functional~\cite{roth2002whitebear}.  The association term is
    taken from~\cite{yu2002fmt-dft-inhomogeneous-associating}, and we
    used a simple and efficient functional for the dispersion
    interaction with a single tunable parameter.
  }

  \headerbox{Surface Tension}{name=1D,column=1,row=0}{
    \begin{center}
      \includegraphics[width=\columnwidth]{figs/surface-tension}
    \end{center}
  }
  \headerbox{A Cylinder in Water}{name=rod,column=1,below=1D,span=1,above=bottom}{
    \begin{center}
      \includegraphics[width=\columnwidth]{figs/density-vs-radius}
    \end{center}
    \begin{center}
      \includegraphics[width=\columnwidth]{figs/energy-vs-diameter}
    \end{center}
  }
  \headerbox{Two Cylinders in Water}{name=two rods,column=2,span=2,row=0}{
    \begin{multicols}{2}
      This text will be in two separate columns.
      This text will be in two separate columns.
      This text will be in two separate columns.
      This text will be in two separate columns.
      This text will be in two separate columns.
      This text will be in two separate columns.
      This text will be in two separate columns.
    \end{multicols}\vspace{-1em}
    \begin{center}
      \includegraphics[width=0.45\columnwidth]{figs/density-rods-in-water}
      \hspace{1em}
      \includegraphics[width=0.45\columnwidth]{figs/xassoc-rods-in-water}
    \end{center}

    %      \begin{multicols}{2}
    {
      This text will span the whole box.
      This text will span the whole box.
      This text will span the whole box.
      This text will span the whole box.
      This text will span the whole box.
      This text will span the whole box.
      %Note the different scales for the two graphs.
      %Our approach has a high accuracy on the neutral (UND) dataset.
    }
    %      \end{multicols}\vspace{-1em}
    \begin{center}
      \includegraphics[width=0.45\columnwidth]{figs/rods-energy-vs-distance}
    \end{center}
  }%
  \headerbox{Future Work}{name=future work,column=2,span=1,above=bottom,below=two rods}{

  }
  \headerbox{References}{name=references,column=3,above=bottom,below=two
  rods}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \bibliographystyle{plain}	% (uses file "plain.bst")
    \bibliography{paper}
    %% \smaller
    %% \vspace{-0.4em}
    %% \bibliographystyle{ieee}
    %% \renewcommand{\section}[2]{\vskip 0.05em}
    %%   \begin{thebibliography}{1}\itemsep=-0.01em
    %%   \setlength{\baselineskip}{0.4em}
    %%   \bibitem{amberg07:nonrigid}
    %%     B.~Amberg, S.~Romdhani, T. Vetter.
    %%     \newblock {O}ptimal {S}tep {N}onrigid {ICP} {A}lgorithms for {S}urface {R}egistration
    %%     \newblock In {\em CVPR 2007}
    %%   \bibitem{amberg08:recognition}
    %%     B.~Amberg, R.~Knothe, T. Vetter.
    %%     \newblock Expression Invariant Face Recognition with a 3D Morphable Model
    %%     \newblock In {\em AFGR 2008}
    %%   \end{thebibliography}
  }%
\end{poster}%
%
\end{document}

>>>>

\definecolor{cyan}{cmyk}{0.4,0,0.3,0}
\definecolor{greenblue}{rgb}{0.7,1,.7}
\definecolor{mediumblue}{rgb}{0.7,0.7,1}
\begin{poster}{
  % Show grid to help with alignment
  grid=no,
  % Column spacing
  colspacing=1em,
  % Color style
  bgColorOne=lightyellow,
  bgColorTwo=lightestyellow,
  borderColor=mediumblue,
  headerColorOne=lighterblue,
  headerColorTwo=mediumblue,
  headerFontColor=black,
  boxColorOne=lighterblue,
  boxColorTwo=lightblue,
  % Format of textbox
  textborder=roundedleft,
  % Format of text header
  eyecatcher=yes,
  headerborder=open,
  headerheight=0.08\textheight,
  headershape=roundedright,
  headershade=shade-tb,
  headerfont=\Large\textsf, %Sans Serif
  boxshade=plain,
%  background=shade-tb,
  background=shade-tb,
  linewidth=2pt
  }
  % Eye Catcher
  { % The title is centered between eye-catcher and logo.
    \includegraphics[height=5.5em]{figs/cavity-cartoon}
  }
  % Title
  {\sf %Sans Serif
  %\bf% Serif
  A Classical Density Functional for Water}
  % Authors
  {\sf %Sans Serif
  % Serif
  David Roundy\hspace{3em}
  Jessica Hughes\hspace{3em}
  Oregon State University, USA
  }
  % University logo
  {
    \includegraphics[height=5.5em]{figs/osu-logo}
  }

@@

  \headerbox{Classical DFT}{name=cdft,column=0,row=0}{ Classical
    Density-Functional Theory is a continuum empirical approach to
    modelling fluid behavior when the density is inhomogeneous---as is
    the case at interfaces.  We have developed a classical density
    functional for water, which provides a description of hydrophobic
    interactions.

    \vspace{1em}
  }

  \headerbox{Statistical Associating Fluid
    Teory}{name=saft,column=0,below=cdft}{ Statistical Associating
    Fluid Theory (SAFT) is a well-established approach for the
    description of the thermodynamics and phase equilibria of fluid
    systems, with particular applicability to hydrogen-bonding fluids.
    Our functional is constructed so as to reduce in the homogeneous
    limit to the 4-site SAFT model for water developed in
    reference~\cite{clark2006developing}.
    \begin{center}
      \includegraphics[width=\columnwidth]{figs/pressure-with-isotherms}
    \end{center}
    This functional accurately describes the liquid-vapor phase
    diagram away from the critical point, but gives less accurately
    describes the compressibility.

    \vspace{1em}
  }

  \headerbox{The Functional}{name=functional,column=1,row=0}{
    The Helmholtz free energy of the new functional is based on SAFT,
    and is composed of the usual terms:
    \begin{equation*}
      F[n] = F_\textit{id}[n] + F_\textit{hs}[n] + F_\textit{assoc}[n] + F_\textit{disp}[n],
    \end{equation*}
    We model the repulsive interactions using the White Bear version
    of the Fundamental-Measure Theory~(FMT)
    functional~\cite{roth2002whitebear} for the hard sphere fluid,
    which is expressed in terms of efficiently-computable
    convolutions.
    \vspace{0.5em}

    The SAFT free energy of association for a 4-site model of water
    has the form
    \begin{align*}
      f_\text{assoc} &= 4 k_BT n \left(\ln X - \frac{X}{2} + \frac12\right)
    \end{align*}
    where $X$ is the fraction of hydrogens \emph{not} hydrogen bonded,
    and is found using a mass action equation involving the density of
    hard spheres in contact with one another.  We extend this to the
    inhomogeneous case using the approach described by Yu and
    Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}, which expresses
    $X$ in terms of the same weighted densities used in FMT.
    \vspace{0.5em}

    Finally, we add a simple and efficient functional for the
    dispersion interaction based on
    SAFT-VR~\cite{gil-villegas-1997-SAFT-VR}, with an additional
    tunable parameter to adjust the surface tension.

    \vspace{0.8em}
  }

  \headerbox{Surface
    Tension}{name=surfacetension,column=1,below=functional,above=bottom}{
    We compute the surface tension by finding the free energy of a
    liquid-vapor interface.  The surface tension is fit to experiment
    by adjusting a length-scale parameter in the dispersion term of
    the free energy.
    \begin{center}
      \includegraphics[width=\columnwidth]{figs/surface-tension}
    \end{center}
  }
  \headerbox{A Single Cylinder in Water}{name=rod,column=2,span=2,row=0}{
    \begin{multicols}{2}
      One of the simplest multi-dimensional inhomogeneous systems is a
      hard cylinder in water.
      \begin{center}
        \includegraphics[width=0.9\columnwidth]{figs/density-vs-radius}
      \end{center}
      \begin{quote}
        The density of water around a
        hard-walled cylinder with 1~nm diameter.
      \end{quote}

      \begin{center}
        \includegraphics[width=\columnwidth]{figs/energy-vs-diameter}
      \end{center}
      \begin{quote}
        The hydration energy per unit area of a hard
        cylinder.  As expected, this asymptotes to the same value as
        the surface tension.
      \end{quote}
    \end{multicols}
  }
  \headerbox{Two Cylinders in Water}{name=two rods,column=2,span=2,below=rod,above=bottom}{
    %% \begin{multicols}{2}
    %%   This text will be in two separate columns.
    %%   This text will be in two separate columns.
    %%   This text will be in two separate columns.
    %%   This text will be in two separate columns.
    %%   This text will be in two separate columns.
    %%   This text will be in two separate columns.
    %%   This text will be in two separate columns.
    %% \end{multicols}\vspace{-1em}
    \begin{multicols}{2}
      \includegraphics[width=\columnwidth]{figs/density-rods-in-water}

      \vspace{-1em}
      \begin{quote}
        The density of water in the vicinity of two hard-walled
        cylinders, each 1~nm in diameter, and separated by 1~nm.
      \end{quote}

      \includegraphics[width=\columnwidth]{figs/xassoc-rods-in-water}

      \vspace{-1em}
      \begin{quote}
      Here we plot the number of hydrogen bonds per molecule for the
      same system of two hard-walled cylinders.
      \end{quote}
    \end{multicols}

    \begin{multicols}{2}
      One application of our model is the hydrophobic attraction
      between two cylindrical molecules in water, such as two
      nanotubes.  We study the interactions between two hard cylinders
      with 1~nm diameter.
      \begin{center}
        \includegraphics[width=0.95\columnwidth]{figs/two-rods-diagram}
      \end{center}
      \begin{center}
        \includegraphics[width=0.9\columnwidth]{figs/rods-energy-vs-distance}
      \end{center}
    \end{multicols}
  }%
  %% \headerbox{Future Work}{name=future work,column=3,span=1,above=two rods,row=0}{
  %%   \begin{itemize}
  %%   \item Extend functional to depend on hydrogen and oxygen densities
  %%     separately, including intramolecular interactions in the
  %%     harmonic approximation.
  %%   \item Introduce elecrtrostatic interactions.
  %%   \item Couple with.
  %%   \end{itemize}
  %% }
  \headerbox{References}{name=references,column=0,above=bottom,below=saft}{
    \renewcommand\refname{}
    \bibliographystyle{unsrt}
    \tiny
    \bibliography{paper}
  }%
\end{poster}%
%
\end{document}

### 2011-06-27 paper/project.tex

@@

>>>>

\textcolor{red}{I found a nice paper on the hard-slit
  problem... actually a couple of papers that address
  it~\cite{forsman1996computer, lum1999hydrophobicity}}

@@

Here we will want to cite~\cite{walther2004hydrodynamic}, which gives
simulations for two interacting nanotubes in water.  It's not
precisely the same, but if we can't find something more similar, this
would be worth citing.
